Computing properties of independent sets by generic programming tensor networks \thanks{\funding{...}}

### 2022-03-15 paper/paper_maddie_edits.tex
added comments on general flow and questions on content

%%%%%% Comment related
\newcommand{\red}[1]{[{\bf  \color{red}{ST: #1}}]}
\newcommand{\xred}[1]{[{\bf  \color{red}{\sout{ST: #1}}}]}
\newcommand{\green}[1]{[{\bf  \color{green}{XG: #1}}]}
\newcommand{\xgreen}[1]{[{\bf  \color{green}{\sout{XG: #1}}}]}
\newcommand{\blue}[1]{[{\bf  \color{blue}{JG: #1}}]}
\newcommand{\xblue}[1]{[{\bf  \color{blue}{\sout{JG: #1}}}]}
\newcommand{\cyan}[1]{[{\bf  \color{cyan}{ML: #1}}]}
\newcommand{\xcyan}[1]{[{\bf  \color{cyan}{\sout{ML: #1}}}]}
\newcommand{\material}[1]{\iffalse[{\bf  \color{cyan}{Material: #1}}]\fi}

@@

\section{Introduction}
In graph theory and combinatorial optimization, there is an important class of problems that can be quantified over a vertex and its neighborhood,
including the independent set problem, the cutting problem, the K-coloring problem, the satisfiability problem,
the maximal clique problem, and the vertex cover problem et al ~\cite{Moore2011}.
There is a widespread interest of these problem because they have wide range of applications in scheduling, logistics,
wireless networks and telecommunication, and computer vision, etc.~\cite{Butenko2003, Wu2015}.
\blue{need more diversed applications not limited to IS}

The most studied version of the above problems is finding a maximum/minimum set satisfying the contraints and the corresponding set size ,
while many of them belong to the hardness category NP-hard~\cite{Hastad1996}, i.e. it is unlikely to solve them in a polynomial time.
In this paper, we focus on the less studied \textit{property} computation which can be harder but is crutial towards understanding a hard problem.
Here, the \textit{property} refers to a class of quantities that not only include the maximum/minimum set size,
but also include the number of sets at a give size, enumeration of all sets at a give size or direct sampling of such sets when they are not enumeratable.
For the weighted version of the above problems, it also includes finding largest/smallest $k$ sets and their sizes.
%are many interesting and hard computational problems concerning various properties of independent sets.

We propose to view the computation of all these properties from an abstract algebra perspective and use a unified framework,
generic tensor-network, to solve these seeming unrelated properties all.
%We map the computation of these properties into generic tensor network contraction with specially designed tensor element algebra.
Tensor network is also known as the sum product network or einsum in different fields,
where its tensor element types are typically restricted to standard number types such as real numbers and complex numbers.
We extend this concept to generic tensor network by generalizing the tensor element type to any type being a commutative semiring.
Given a good contraction order, the space required to contract a tensor network is 2 power the tree width of its line graph~\cite{Markov2008}.
Recent progress in simulating quantum circuits with random tensor networks~\cite{Gray2021, Pan2021, Kalachev2021} enables us to find such a good contraction order for a tensor network with thousands of tensors in a reasonable time.
When using tensor network for solving one of the best sets with the maximum size,
this time complexity is just on par with the dynamic programming algorithms~\cite{Courcelle1990, Fomin2013},
however, tensor network has richer algebraic structure, and this structure is crutial for property computation.
%For sparse graphs, the treewidth is usually much smaller than the number of vertices~\cite{Fomin2006}. Our algorithms, on the other hand, are much more versatile than traditional methods and can be used to compute many other properties than just the MIS size.

@@

To avoid the discussion being unnecessaryly diversed, we mainly focus on the independent set problem,
and discuss other combinatorial optimization problems in \App{sec:otherproblems}.
The paper is organized as follows.
We first introduce the basic concepts of tensor networks and generic programming in \Sec{sec:tn} and \Sec{sec:generic}.
Then we show how to reduce an independent set problem to a tensor network contraction problem \Sec{sec:tnmap}
and show how to engineer the element types in \Sec{sec:counting},\Sec{sec:enumeration} and \Sec{sec:weighted} to count independent sets at a given size, enumerate/sample independent sets at a given size and find largest set sizes in weighted graphs.
% by computing a number of independent set properties of certain sparse graphs on central processing units (CPUs) and graphics processing units (GPUs) and show its good performance.
Lastly, we benchmark our algorithms in \Sec{sec:benchmark} and provide a few examples in \Sec{sec:examples} to demonstrate the versatility of our tool.
We show how our method can help computing the entropy constant for some hardcore lattice gases on 2D square lattices
and analyzing the presence of overlap gap property on three regular graphs.
%Our method can also be used to find ``maximal'' independent set properties; 
%a maximal independent set is an independent set that is not a subset of any other independent set, but its size may not be the maximum.

\section{Tensor networks}\label{sec:tn}
Tensor network~\cite{Cirac2020, Orus2014} is also known as einsum, factor graph or sum-product network~\cite{Bishop2006} in different contexts;
it can be viewed as a generalization of binary matrix multiplication to nary tensor contraction.
Einstein's notation is often used to represent a tensor network, e.g.\ it represents the matrix multiplication between two matrices $A$ and $B$ as $C_{ik} = A_{ij}B_{jk}$,
where we use a label in the subscript to represent a degree of freedom.
One can enumerate these degrees of freedom and accumulate the product of tensor elements to the output tensor.
In the standard Einstein's notation for tensor networks in physics, each index appears precisely twice.
Hence a tensor network can be represented as a simple graph,
where a tensor is mapped to a vertex and a label shared by two tensors is mapped to an edge.
In this work, we do not impose such a restriction, so an index can appear an arbitrary number of times. 
The graphical representation of a generalized tensor network is a hypergraph, in which an edge (a label in subscript) can be shared by an arbitrary number of vertices (tensors).
These two notations are equivalent in representation power because one can easily translate a generalized tensor network to the standard notation by adding $\delta$ tensors, a high dimensional equivalence of the identity matrix.
However, introducing $\delta$ tensors may significantly increase the complexity of the tensor contraction. We illustrate this subtle point in \App{sec:tensorbad}.

@@

\section{Generic programming tensor contractions}\label{sec:generic}
In previous works relating tensor networks and combinatoric problems~\cite{Kourtis2019, Biamonte2017},
the elements in the tensor networks are limited to standard number types such as floating point numbers and integers.
Owning to the development of modern compiling technology, we no longer need to re our imagination to standard number types.
One of the key concepts that push the technology forward is called generic programming:

\begin{definition}[Generic programming ~\cite{Stepanov2014}]
   Generic programming is an approach to programming that focuses on designing algorithms and data structures so that they work in the most general setting without loss of efficiency.
\end{definition}

This definition of generic programming contains two major aspects: a single program works in the most general setting and efficiency.
To understand the first aspect on generality, suppose we want to write a function that raises an element to a power, $f(x, n) := x^n$.
One can easily write a function for standard number types that computes the power of $x$ in $O \left( \log(n) \right)$ steps using the multiply and square trick.
Generic programming does not require $x$ to be a standard number type,
instead it treats $x$ as an element with an associative multiplication operation $\odot$ and a multiplicative identity $\mymathbb{1}$.
In such a way, when the program takes a matrix as an input, it computes the matrix power without extra efforts.
The second aspect is about performance. For dynamically typed languages such as Python,
one can easily write very general codes, but the efficiency is not guaranteed; for example, the speed of computing the matrix multiplication between two numpy arrays with python objects as elements is much slower than statically typed languages such as C++ and Julia~\cite{Bezanson2012}.
C++ uses templates for generic programming while Julia takes advantage of just-in-time compilation and multiple dispatch.
When these languages ``see'' a new input type, the compiler can recompile the generic program for the new type.
A myriad of optimizations can be done during the compilation, such as inlining immutable elements with fixed sizes in an array to decrease the cache miss rate when accessing data.
In Julia, these inlined arrays can even be compiled to GPU devices for faster computation~\cite{Besard2018}.

This motivates us to think about what is the most general element type allowed in a tensor network contraction program.
We find that as long as the algebra of tensor elements forms a commutative semiring, the tensor network contraction result will be valid and be independent of the contraction order.
A commutative semiring is a ring with its multiplication operation being commutative and without an additive inverse.
To define a commutative semiring with the addition operation $\oplus$ and the multiplication operation $\odot$ on a set $R$, the following relations must hold for any arbitrary three elements $a, b, c \in R$.
\begin{align*}
(a \oplus b) \oplus c = a \oplus (b \oplus c) & \hspace{5em}\text{$\triangleright$ commutative monoid $\oplus$ with identity $\mymathbb{0}$}\\
a \oplus \mymathbb{0} = \mymathbb{0} \oplus a = a &\\
a \oplus b = b \oplus a &\\
&\\
(a \odot b) \odot c = a \odot (b \odot c)  &   \hspace{5em}\text{$\triangleright$ commutative monoid $\odot$ with identity $\mymathbb{1}$}\\
a \odot  \mymathbb{1} =  \mymathbb{1} \odot a = a &\\
a \odot b = b \odot a &\\
&\\
a \odot (b\oplus c) = a\odot b \oplus a\odot c  &  \hspace{5em}\text{$\triangleright$ left and right distributive}\\
(a\oplus b) \odot c = a\odot c \oplus b\odot c &\\
&\\
a \odot \mymathbb{0} = \mymathbb{0} \odot a = \mymathbb{0}
\end{align*}
We require the algebra being multiplicative commutative because the contraction order optimizer will change the multiplication order of elements, while we wish this contraction order does not change the contraction results.
In the following sections, we show how to compute a number of properties of independent sets by designing tensor element types as specific commutative semirings without changing the tensor network contraction program~\cite{Stepanov2014}.
The Table~\ref{tbl:generictypes} summarizes those properties that can be computed by various tensor element types.

\begin{table}[t!]\centering
\begin{minipage}{\columnwidth}
\ra{1.3}
        \begin{tabularx}{\textwidth}{sb}\toprule
            \hline
   \textbf{Element type}     & \textbf{Property to compute} \\
   {$\mathbbm{R}$}     & {Counting of all independent sets} \\
   {Polynomial} (\Eq{eq:polynomial})     & {Independence polynomial} \\
   {Tropical (\Eq{eq:tropical})}    & {Maximum independent set size} \\
   {Extended tropical of order $k$ (\Eq{eq:ext-tropical})}    & {Largest $k$ independent set sizes} \\
   {Polynomial truncated to $k$-th order (\Eq{eq:countingtropical} and \Eq{eq:max2poly})}     & {$k$ largest independent sizes and their counting} \\
   {Set} (\Eq{eq:set})     & {Enumeration of independent sets} \\
   {Sum Product Tree} (\Eq{eq:expr})     & {Sampling of independent sets} \\
   {Polynomial truncated to highest order combined with Bit string} (\Eq{eq:singleconfig})     & {Maximum independent set size and one of such configurations} \\
   {Polynomial truncated to $k$-th order combined with Set} (\Eq{eq:countingtropicalset})    & {$k$ largest independent set sizes and their enumeration} \\
            \bottomrule
        \end{tabularx}
    \caption{Tensor element types and the independent set properties that can be computed using them.}\label{tbl:generictypes}
\end{minipage}
\end{table}

\section{Tensor network representation of independent sets} \label{sec:tnmap}
This section is about reducing the independent set problem to a tensor network contraction problem.
For people with physics background, we suggest using an alternative approach described in \App{sec:energymodel} to understand this reduction from an energy model.
Let $G=(V,E)$ be a weighted undirected graph with each vertex $i\in V$ being associated with a weight $w_i$.
An independent set $I \subseteq V$ is a set of vertices that for any vertex pair $u,v \in I$, $(u, v) \not\in E$,
and we denote the maximum independent set (MIS) size $\alpha(G) \equiv \max_{I}\sum_{v\in I}w_v$. 
We map a vertex $v\in V$ to a label $s_v \in \{0, 1\}$ of dimension $2$ in a tensor network, where we use $0$ ($1$) to denote a vertex is absent (present) in the set.
For each label $s_v$, we defined a parametrized rank-one vertex tensor $W(x_v, w_v)$ as
\begin{equation}
    W(x_v, w_v) = \left(\begin{matrix}
        1_v \\
        x_v^{w_v}
    \end{matrix}\right).\label{eq:vertextensor}
\end{equation}
where we use subscripts to annotate different vertices, these annotation are only used in configuration enumeration and sampling.
One can index tensor elements by subscripting tensors, e.g.\ $W(x_i, w_i)_0=1$ is the first element associated with $s_i=0$ and $W(x_i, w_i)_1=x_i^{w_i}$ is the second element associated with $s_i=1$.
Similarly, on each edge $(u, v)$, we define a matrix $B$ indexed by $s_u$ and $s_v$ as
\begin{equation}
    \qquad \quad 
       B = \left(\begin{matrix}
        1  & 1\\
        1 & 0
    \end{matrix}\right). \label{eq:edgetensor}
\end{equation}
The corresponding tensor network contraction gives
\begin{equation}\label{eq:idp}
    P(G, \{x_1^{w_1}, x_{2}^{w_2}, \ldots,x_{|V|}^{w_{|V|}}\}) = \sum\limits_{s_1, s_2, \ldots, s_{|V|} = 0}^{1} \prod\limits_{i=1}^{|V|} W(x_i, w_i)_{s_i} \prod\limits_{(i,j) \in E(G)} B_{s_i s_j},
\end{equation}
where the summation runs over all vertex configurations $\{s_1, s_{2}, \ldots,s_{|V|}\}$ and accumulates the product of tensor elements to the output $P$ (see Example \ref{eg:tensorcontraction} for a concrete example).
The edge tensor element $B_{s_{i}=1, s_{j}=1} = 0$ encodes the independent set constraint, meaning vertex $i$ and $j$ cannot be both in the independent set if they are connected by an edge $(i,j)$.

\begin{example}\label{eg:twonode}
In the following, we show a minimum example of mapping the independent problem of a 2 vertex complete graph K2 (left) to a tensor network (right).

@@

In the graphical representation of tensor network on the right panel,
we use a circle to represent a tensor, a hyperedge in cyan color to represent the degree of freedom $s_a$,
and a hyperedge in red color to represent a degree of freedom $s_b$.
Tensors sharing a same degree of freedom are connected by that hyperedge.
The contraction of this tensor network can be evaluated manually:
\begin{equation}
    \begin{split}
    P({\rm K2},\{x_a^{w_a}, x_b^{w_b}\}) &=
    \left(\begin{matrix} 1 & x_a^{w_a} \end{matrix}\right)
    \left(\begin{matrix} 1 & 1\\ 1  & 0 \end{matrix}\right)
    \left(\begin{matrix} 1 \\ x_b^{w_b} \end{matrix}\right)\\
    &= 1 + x_a^{w_a} + x_b^{w_b}
    \end{split}
\end{equation}
The resulting polynomial represents 3 different independent sets $\{\}$, $\{a\}$ and $\{b\}$, with weights being $0$, $w_a$ and $w_b$ respectively.
\end{example}

It is in general not computational efficient to evaluate \Eq{eq:idp} by directly summing up the $2^{|V|}$ products.
The standard approach to evaluate a tensor network is to contract two tensors at a time with a certain order utilizing the associativity and commutativity of tensor contraction.
A fully optimized contraction order can reduce the time to $O(2^{\text{tw}(G)})$~\cite{Markov2008}, while requiring a space at the same order of magnitude to store intermediate contraction results.
The pairwise tensor contraction also makes it possible to utilize basic linear algebra subprograms (BLAS) functions to speed up the computation for certain tensor element types.

@@

One can easily check the maximum tensor during computing is rank 2, and the total number of multiplication operations is reduced significantly.
\end{example}

>>>>

%%%%%% Comment related
\newcommand{\red}[1]{[{\bf  \color{red}{ST: #1}}]}
\newcommand{\xred}[1]{[{\bf  \color{red}{\sout{ST: #1}}}]}
\newcommand{\green}[1]{[{\bf  \color{green}{XG: #1}}]}
\newcommand{\xgreen}[1]{[{\bf  \color{green}{\sout{XG: #1}}}]}
\newcommand{\blue}[1]{[{\bf  \color{blue}{JG: #1}}]}
\newcommand{\xblue}[1]{[{\bf  \color{blue}{\sout{JG: #1}}}]}
\newcommand{\cyan}[1]{[{\bf  \color{cyan}{ML: #1}}]}
\newcommand{\xcyan}[1]{[{\bf  \color{cyan}{\sout{ML: #1}}}]}
\newcommand{\purple}[1]{[{\bf  \color{purple}{MC: #1}}]}
\newcommand{\xpurple}[1]{[{\bf  \color{purple}{\sout{MC: #1}}}]}
\newcommand{\material}[1]{\iffalse[{\bf  \color{cyan}{Material: #1}}]\fi}

@@

\section{Introduction}
In graph theory and combinatorial optimization, there is an important class of problems that can be quantified over a vertex and its neighborhood,
including the independent set problem, the cutting problem, the K-coloring problem, the satisfiability problem,
the maximal clique problem, and the vertex cover problem et al ~\cite{Moore2011}. \purple{Not clear to me what ``over a vertex and its neighborhood'' means. }
There is a widespread interest of these problem because they have wide range of applications in scheduling, logistics,
wireless networks and telecommunication, and computer vision, etc.~\cite{Butenko2003, Wu2015}.
\blue{need more diversed applications not limited to IS} \purple{Here I would say that that these problems are intractable (exponential runtime in problem size in the worst-case), but it is nevertheless desirable to have high-performance algorithms because of their ubiquitous applications.}

The most studied version of the above problems is finding a maximum/minimum set satisfying the constraints and the corresponding set size,
while many of them belong to the hardness category NP-hard~\cite{Hastad1996}, i.e. it is unlikely to solve them in a polynomial time. \purple{Is MIS really the most studied NP-hard optimization problem? I would guess it is heavily studied, but not necessarily the most studied.}
In this paper, we focus on the less studied \textit{property} computation which can be harder but is crucial towards understanding a hard problem. \purple{It might be interesting to note that computing the number of sets of a given size is \#P-hard (class of all counting problems), which is believed to be much harder than NP.}
Here, the \textit{property} \purple{Consider using ``solution space property'' instead of just ``property''.} refers to a class of quantities that not only include the maximum/minimum set size,
but also include the number of sets at a give size, enumeration of all sets at a give size or direct sampling of such sets when they are not enumeratable.
For the weighted version of the above problems, it also includes finding largest/smallest $k$ sets and their sizes. \purple{I would give examples of why we want to compute solution space properties. For instance, in statistical physics, computing thermal states and observables via the partition function. In complexity theory, proving limits on local algorithms (OGP). }
%are many interesting and hard computational problems concerning various properties of independent sets.

We propose to view the computation of all these properties from an abstract algebra perspective and use a unified framework,
generic tensor-network, to solve these seeming unrelated properties all.
%We map the computation of these properties into generic tensor network contraction with specially designed tensor element algebra.
Tensor network is also known as the sum product network or einsum in different fields, \purple{Say which fields}
where its tensor element types are typically restricted to standard number types such as real numbers and complex numbers. \purple{Briefly say what tensor networks are in a few words, so that ``restricted to standard number types such as real numbers and complex numbers'' can be put into context. A reader new to tensor networks might be wondering what a tensor network is, and the only context we get from this sentence is that they involve numbers}
We extend this concept to generic tensor network by generalizing the tensor element type to any type being a commutative semiring.
Given a good contraction order, the space required to contract a tensor network is 2 power the tree width of its line graph~\cite{Markov2008}. \purple{I'd write out $2^{\text{tw}(G)}$.}
Recent progress in simulating quantum circuits with random tensor networks~\cite{Gray2021, Pan2021, Kalachev2021} enables us to find such a good contraction order for a tensor network with thousands of tensors in a reasonable time. \purple{Is the scaling of these algorithms known?}
When using tensor network for solving one of the best sets with the maximum size,
this time complexity is on par with the dynamic programming algorithms~\cite{Courcelle1990, Fomin2013},
however, tensor network has richer algebraic structure, and this structure is crucial for property computation. \purple{Not clear how first part and second part of sentence are related}
%For sparse graphs, the treewidth is usually much smaller than the number of vertices~\cite{Fomin2006}. Our algorithms, on the other hand, are much more versatile than traditional methods and can be used to compute many other properties than just the MIS size.

@@

To avoid the discussion being unnecessarily complicated, we  focus concretely on the independent set problem,
and discuss other combinatorial optimization problems in \App{sec:otherproblems}. \purple{More details on which combinatorial optimization problems this applies to?}
The paper is organized as follows.
We first introduce the basic concepts of tensor networks and generic programming in \Sec{sec:tn} and \Sec{sec:generic}.
Then we show how to reduce an independent set problem to a tensor network contraction problem \Sec{sec:tnmap}
and show how to engineer the element types in \Sec{sec:counting},\Sec{sec:enumeration} and \Sec{sec:weighted} to count independent sets at a given size, enumerate/sample independent sets at a given size and find largest set sizes in weighted graphs.
% by computing a number of independent set properties of certain sparse graphs on central processing units (CPUs) and graphics processing units (GPUs) and show its good performance.
Lastly, we benchmark our algorithms in \Sec{sec:benchmark} and provide a few examples in \Sec{sec:examples} to demonstrate the versatility of our tool.
We show how our method can help computing the entropy constant for some hardcore lattice gases on 2D square lattices
and analyzing the presence of overlap gap property on three regular graphs.
%Our method can also be used to find ``maximal'' independent set properties; 
%a maximal independent set is an independent set that is not a subset of any other independent set, but its size may not be the maximum.

\section{Tensor networks}\label{sec:tn}
Tensor network~\cite{Cirac2020, Orus2014} is also known as einsum, factor graph or sum-product network~\cite{Bishop2006} in different contexts;
it can be viewed as a generalization of binary matrix multiplication to nary tensor contraction.
Einstein's notation is often used to represent a tensor network, e.g.\ it represents the matrix multiplication between two matrices $A$ and $B$ as $C_{ik} = A_{ij}B_{jk}$,
where we use a label in the subscript to represent a degree of freedom.
One can enumerate these degrees of freedom and accumulate the product of tensor elements to the output tensor.
In the standard Einstein's notation for tensor networks in physics, each index appears precisely twice \purple{signifying which pairs of indices are summed over}.
Hence a tensor network can be represented as a simple graph,
where a tensor is mapped to a vertex and a label shared by two tensors is mapped to an edge.
In this work, we do not impose such a restriction, so an index can appear an arbitrary number of times. 
The graphical representation of a generalized tensor network is a hypergraph, in which an edge (a label in subscript) can be shared by an arbitrary number of vertices (tensors).
These two notations are equivalent in representation power because one can easily translate a generalized tensor network to the standard notation by adding $\delta$ tensors, a high dimensional equivalence of the identity matrix.
However, introducing $\delta$ tensors may significantly increase the complexity of the tensor contraction. We illustrate this subtle point in \App{sec:tensorbad}. \purple{ I feel that the $\delta$ tensors point is unnecessary to mention in the main text. Because not much detail is given, I internalized what it meant only on the third read-through.}

@@

\section{Generic programming tensor contractions}\label{sec:generic}
\purple{On a second read of this section, I think that the clarity could be improved. My understanding is this is the flow: 
We would like to use generic programming, where we write code that works regardless of input type, and is optimized on all input types without sacrificing efficiency. 
In general, we can write code such that a function runs correctly regardless of the input data type. This is easy to do for dynamically typed languages like Python, which only verify and enforce the rules for the input data type at runtime (called \textit{type-checking}). However, when the input data type is only checked at runtime (as opposed to when the code is compiled), type-specific optimization cannot be used. High efficiency  is only achieved in statically typed languages, where modern compiling technology allows the function to be individually optimized for each input data type (then give C++/Julia examples). }
In previous works relating tensor networks and combinatoric problems~\cite{Kourtis2019, Biamonte2017},
the elements in the tensor networks are limited to standard number types such as floating point numbers and integers.
Owing to the development of modern compiling technology, we no longer need to restrict our imagination to standard number types.
One of the key concepts that push the technology forward is called generic programming:

\begin{definition}[Generic programming ~\cite{Stepanov2014}]
   Generic programming is an approach to programming that focuses on designing algorithms and data structures so that they work in the most general setting without loss of efficiency.
\end{definition}
\purple{I think the following paragraph could be clarified. When you read the definition of generic programming, it's not clear what ``in the most general setting'' means. My understanding is it just means that the program will work regardless of the input data type - it would be good to say this explicitly. I also think the way it is written takes a while to understand what dynamically versus statically typed languages are. The definition, ``When these languages ``see'' a new input type, the compiler can recompile the generic program for the new type.'' comes rather late in the paragraph. Also, while I know what you mean, concepts like ``inlining immutable elements'' and ``cache miss rate'' would be lost on someone only with a physics background. }

This definition of generic programming contains two major aspects: a single program works in the most general setting and efficiency.
To understand the first aspect on generality, \purple{This is a really helpful example!} suppose we want to write a function that raises an element to a power, $f(x, n) := x^n$.
One can easily write a function for standard number types that computes the power of $x$ in $O \left( \log(n) \right)$ steps using the multiply and square trick.
Generic programming does not require $x$ to be a standard number type,
instead it treats $x$ as an element with an associative multiplication operation $\odot$ and a multiplicative identity $\mymathbb{1}$.
In such a way, when the program takes a matrix as an input, it computes the matrix power without extra effort.
The second aspect is about performance. For dynamically typed languages such as Python,
one can easily write very general codes, but the efficiency is not guaranteed; for example, the speed of computing the matrix multiplication between two numpy arrays with python objects as elements is much slower than statically typed languages such as C++ and Julia~\cite{Bezanson2012}.
C++ uses templates for generic programming while Julia takes advantage of just-in-time compilation and multiple dispatch.
When these languages ``see'' a new input type, the compiler can recompile the generic program for the new type.
A myriad of optimizations can be done during the compilation, such as inlining immutable elements with fixed sizes in an array to decrease the cache miss rate when accessing data.
In Julia, these inlined arrays can even be compiled to GPU devices for faster computation~\cite{Besard2018}.

This motivates us to think about what is the most general element type \purple{I would clarify element type means tensor element type} allowed in a tensor network contraction program.
We find that as long as the algebra of tensor elements forms a commutative semiring, the tensor network contraction result will be valid and be independent of the contraction order.
A commutative semiring is a ring with its multiplication operation being commutative and without an additive inverse.
To define a commutative semiring with the addition operation $\oplus$ and the multiplication operation $\odot$ on a set $R$, the following relations must hold for any arbitrary three elements $a, b, c \in R$.
\begin{align*}
(a \oplus b) \oplus c = a \oplus (b \oplus c) & \hspace{5em}\text{$\triangleright$ commutative monoid $\oplus$ with identity $\mymathbb{0}$}\\
a \oplus \mymathbb{0} = \mymathbb{0} \oplus a = a &\\
a \oplus b = b \oplus a &\\
&\\
(a \odot b) \odot c = a \odot (b \odot c)  &   \hspace{5em}\text{$\triangleright$ commutative monoid $\odot$ with identity $\mymathbb{1}$}\\
a \odot  \mymathbb{1} =  \mymathbb{1} \odot a = a &\\
a \odot b = b \odot a &\\
&\\
a \odot (b\oplus c) = a\odot b \oplus a\odot c  &  \hspace{5em}\text{$\triangleright$ left and right distributive}\\
(a\oplus b) \odot c = a\odot c \oplus b\odot c &\\
&\\
a \odot \mymathbb{0} = \mymathbb{0} \odot a = \mymathbb{0}
\end{align*}
We require the algebra be multiplicative commutative because the contraction order optimizer will change the multiplication order of elements, while we wish this contraction order does not change the contraction results. \purple{The previous sentence makes a lot of sense!}
In the following sections, we show how to compute a number of properties of independent sets by designing tensor element types as specific commutative semirings without changing the tensor network contraction program~\cite{Stepanov2014}.
The Table~\ref{tbl:generictypes} summarizes those properties that can be computed by various tensor element types.

\begin{table}[t!]\centering
\begin{minipage}{\columnwidth}
\ra{1.3}
        \begin{tabularx}{\textwidth}{sb}\toprule
            \hline
   \textbf{Element type}     & \textbf{Property to compute} \\
   {$\mathbbm{R}$}     & {Counting of all independent sets} \\
   {Polynomial} (\Eq{eq:polynomial})     & {Independence polynomial} \\
   {Tropical (\Eq{eq:tropical})}    & {Maximum independent set size} \\
   {Extended tropical of order $k$ (\Eq{eq:ext-tropical})}    & {Largest $k$ independent set sizes} \\
   {Polynomial truncated to $k$-th order (\Eq{eq:countingtropical} and \Eq{eq:max2poly})}     & {$k$ largest independent set sizes and their counting \purple{``and their counting'' is gramatically confusing, can we say independence polynomial or degeneracy or something?}} \\
   {Set} (\Eq{eq:set})     & {Enumeration of independent sets} \\
   {Sum Product Tree} (\Eq{eq:expr})     & {Sampling of independent sets} \\
   {Polynomial truncated to highest order combined with Bit string} (\Eq{eq:singleconfig})     & {Maximum independent set size and one of such configurations} \\
   {Polynomial truncated to $k$-th order combined with Set} (\Eq{eq:countingtropicalset})    & {$k$ largest independent set sizes and their enumeration} \\
            \bottomrule
        \end{tabularx}
    \caption{Tensor element types and the independent set properties that can be computed using them.}\label{tbl:generictypes}
\end{minipage}
\end{table}

\section{Tensor network representation of independent sets} \label{sec:tnmap}
This section is about reducing the independent set problem to a tensor network contraction problem.
For people with a physics background, we suggest using an alternative approach described in \App{sec:energymodel} to understand this reduction from an energy model.
Let $G=(V,E)$ be a weighted, undirected graph with each vertex $i\in V$ being associated with a weight $w_i$ \purple{I suggest being consistent in subscripts, and use $u/v$ for nodes as you do later, e.g. $v \in V$ instead of $i/j$}
An independent set $I \subseteq V$ is a set of vertices that for any vertex pair $u,v \in I$, $(u, v) \not\in E$,
and we denote the maximum independent set (MIS) size $\alpha(G) \equiv \max_{I}\sum_{v\in I}w_v$. 
We map a vertex $v\in V$ to a label $s_v \in \{0, 1\}$ of dimension $2$ in a tensor network, where we use $0$ ($1$) to denote a vertex is absent (present) in the set.
For each label $s_v$, we defined a parametrized rank-one vertex tensor $W(x_v, w_v)$ as
\begin{equation}
    W(x_v, w_v) = \left(\begin{matrix}
        1_v \\
        x_v^{w_v}
    \end{matrix}\right).\label{eq:vertextensor}
\end{equation}
where we use subscripts to annotate different vertices, these annotation are only used in configuration enumeration and sampling.
One can index tensor elements by subscripting tensors, e.g.\ $W(x_i, w_i)_0=1$ is the first element associated with $s_i=0$ and $W(x_i, w_i)_1=x_i^{w_i}$ is the second element associated with $s_i=1$.
Similarly, on each edge $(u, v)$, we define a matrix $B$ indexed by $s_u$ and $s_v$ as
\begin{equation}
    \qquad \quad 
       B = \left(\begin{matrix}
        1  & 1\\
        1 & 0
    \end{matrix}\right). \label{eq:edgetensor}
\end{equation}
The corresponding tensor network contraction gives
\begin{equation}\label{eq:idp}
    P(G, \{x_1^{w_1}, x_{2}^{w_2}, \ldots,x_{|V|}^{w_{|V|}}\}) = \sum\limits_{s_1, s_2, \ldots, s_{|V|} = 0}^{1} \prod\limits_{i=1}^{|V|} W(x_i, w_i)_{s_i} \prod\limits_{(i,j) \in E(G)} B_{s_i s_j},
\end{equation}
where the summation runs over all vertex configurations $\{s_1, s_{2}, \ldots,s_{|V|}\}$ and accumulates the product of tensor elements to the output $P$ (see Example \ref{eg:tensorcontraction} for a concrete example).
The edge tensor element $B_{s_{i}=1, s_{j}=1} = 0$ encodes the independent set constraint, meaning vertex $i$ and $j$ cannot be both in the independent set if they are connected by an edge $(i,j)$.

\begin{example}\label{eg:twonode}
In the following, we show a minimum example of mapping the independent problem of a complete graphs with two vertices K2 (left) to a tensor network (right).

@@

In the graphical representation of tensor network on the right panel,
we use a circle to represent a tensor, a hyperedge in cyan color to represent the degree of freedom $s_a$,
and a hyperedge in red color to represent a degree of freedom $s_b$.
Tensors sharing a same degree of freedom are connected by that hyperedge.
The contraction of this tensor network can be evaluated manually:
\begin{equation}
    \begin{split}
    P({\rm K2},\{x_a^{w_a}, x_b^{w_b}\}) &=
    \left(\begin{matrix} 1 & x_a^{w_a} \end{matrix}\right)
    \left(\begin{matrix} 1 & 1\\ 1  & 0 \end{matrix}\right)
    \left(\begin{matrix} 1 \\ x_b^{w_b} \end{matrix}\right)\\
    &= 1 + x_a^{w_a} + x_b^{w_b}
    \end{split}
\end{equation}
The resulting polynomial represents 3 different independent sets $\{\}$, $\{a\}$ and $\{b\}$, with weights being $0$, $w_a$ and $w_b$ respectively.
\end{example}
\purple{This example is really helpful!}

It is in general not computational efficient to evaluate \Eq{eq:idp} by directly summing up the $2^{|V|}$ products. \purple{Perhaps explicitly say what this scales as?}
The standard approach to evaluate a tensor network is to contract two tensors at a time with a certain order utilizing the associativity and commutativity of tensor contraction.
A fully optimized contraction order can reduce the time to $O(2^{\text{tw}(G)})$~\cite{Markov2008}, while requiring a space at the same order of magnitude to store intermediate contraction results. \purple{Perhaps just say time and space complexity? Order of magnitude doesn't make sense to me in terms of scaling. Also, what is the time complexity of finding a good contraction order? Is there a known average-case complexity? This seems important.}
The pairwise tensor contraction also makes it possible to utilize basic linear algebra subprograms (BLAS) functions to speed up the computation for certain tensor element types.

@@

One can easily check the maximum tensor during computing is rank 2, and the total number of multiplication operations is reduced significantly. \purple{I have not checked this yet but will tomorrow. Is it really easy/simple enough to do without writing stuff down? Is this the simplest example possible? The equations look quite formidable at first glance, and it takes a lot of time to match the hyperedge labels in the equation to the physical graph.}
\end{example}

\section{Independence polynomial}\label{sec:indpoly}
Let $x_i = x$ and $w_i = 1$, the evaluation of \Eq{eq:idp} corresponds to the independence polynomial~\cite{Harvey2018,Ferrin2014}.
which is a useful graph characteristic related to, for example, the partition functions~\cite{Lee1952,Yang1952} and Euler characteristics of the independence complex~\cite{Bousquet2008, Levit2009}.
The independence polynomial is an important graph polynomial that contains the counting information of independent sets. It is defined as
\begin{equation}\label{eq:idpdef}
I(G, x) = \sum_{k=0}^{\alpha(G)} a_k x^k,
\end{equation}
where $a_k$ is the number of independent sets of size $k$ in graph $G=(V,E)$. The total number of independent sets is thus equal to $I(G, 1)$.
Its connection to \Eq{eq:idp} can be understood as follows: the product over vertex tensor elements produces a factor $x^k$, where $k=\sum_i s_i$ counts the set size,
and the product over edge tensor elements gives a factor $1$ for a configuration being in an independent set and $0$ otherwise. The summation counts the number of independent sets of size $k$. 
One can feed the standard computer numbers types into this tensor network and evaluate the independence polynomial by contracting it numerically. \purple{By feed in standard computer number types into the network, do you mean for $x$? Then, you look at multiple $x$, you can fit the independence polynomial to solve for the coefficients? It's not very explicit this procedure works at this point.}
However, we are more interested in finding the coefficients of this polynomial, because this quantity tells us the counting of independent sets at each size. \purple{Does the previous technique (contracting the tensor network numerically) give the coefficients of the polynomial? The use of the word ``however'' here is confusing to me because it feels like you're saying that the previous technique doesn't give the coefficients, or it's very hard to obtain the coefficients.}
To achieve this, let us first elevate the tensor elements $0$s and $1$s in tensors $W(x, 1)$ and $B$ from integers and floating point numbers to the additive identity,
